Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}

### 2020-06-01 Chapters/Conclusions.tex
update 20200602

  	\item Autoregressive B-Format Late IR Estimation \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Novel technique for RT60 estimation from autoregressive models (subproduct of dereverberation)
		\item \textbf{Signal enhancement}: Novel methodology to re-reverberate sound scenes (include new elements in the scene using the reverb of the recorded scene)
	\end{itemize}

  	\item A hybrid parametric-deep learning approach for sound event localization and detection \todo{ref}
  	\begin{itemize}
  		\item \textbf{Scene Description}: ovel State-of-the-Art methodology for Sound Event Localization and Detection	
  	\end{itemize}

@@

		\item Multichannel Array Signal Processing library \todo{ref}
		\begin{itemize}
  			\item \textbf{Simulated IRs}: Library for acoustic simulation (IR generation, microphone array simulation, etc)
		\end{itemize}

  		\item pysofaconventions \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: implementation of SOFA for python 
		\end{itemize}

>>>>

  	\item Blind reverberation time estimation from ambisonic recordings \todo{ref}
  	\begin{itemize}
  		\item \textbf{Parameter estimation}: Novel technique for blind RT60 estimation of ambisonic recordings from autoregressive models
	\end{itemize}

  	\item A hybrid parametric-deep learning approach for sound event localization and detection \todo{ref}
  	\begin{itemize}
  		\item \textbf{Scene Description}: Novel state-of-the-Art methodology for Sound Event Localization and Detection	
  	\end{itemize}

@@

		\item A Python library for Multichannel Acoustic Signal Processing \todo{ref}
		\begin{itemize}
  			\item \textbf{Simulated IRs}: Library for acoustic simulation (IR generation, microphone array simulation, etc)
		\end{itemize}

  		\item pysofaconventions, a Python API for SOFA \todo{ref}
		\begin{itemize}
  			\item \textbf{Recorded IRs}: implementation of SOFA for python 
		\end{itemize}

### 2020-06-01 Chapters/ScientificBackground.tex
update 20200602

\chapter{Scientific Background}

\section{Reference system and conventions}

@@

@@

\section{Ambisonics}

@@

The ordered set of values of all spherical harmonics up to order $N$, evaluated at the source position, is known as the \textit{ambisonic coefficients}:
\begin{equation}
	\pmb{Y}_n^m(\Omega_s) = [Y_0^0(\Omega_s), Y_1^{-1}(\Omega_s),  \ldots ,  Y_N^N(\Omega_s)]
	\label{eq:sphericalharmonicvector}
\end{equation}

Furthermore, the process of multiplying the signal $S$ by the ambisonic coefficients is known in the literature as the \textit{ambisonic encoding}. The resulting signal vector is usually referred to as the \textit{ambisonic} (or \textit{B-Format}) signal $\pmb{S}_n^m$:
\begin{equation}
	\pmb{S}_n^m = \pmb{Y}_n^m(\Omega_s) S
\end{equation}

@@

\todo{change L for other name}
Let us consider a loudspeaker located at the angular position $\Omega_\ell$. The signal feed $L$ is \textit{decoded} from the ambisonic signal as:
\begin{equation}
	L = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) Y_n^m(\Omega_\ell)  S \alpha_n 
	\label{eq:decoding}
\end{equation}

where $\alpha_n$ is a weighting factor which accounts for the beam directivity. Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase} -- the reader is referred to [\todo{find ref. daniel? zotter?}] for more information.

The decoding equation \ref{eq:decoding} can be written in vector form as:
\begin{equation}
	L =  \pmb{S}_n^m {\pmb{Y}_n^{m}(\Omega_\ell)}^T \alpha_n
\end{equation}

where the superscript $T$ represents the matrix transposition. 
This equation can be extended to the regular case of decoding to a loudspeaker array, comprised of $L$ loudspeakers located at the positions $\pmb{\Omega}_l = [\Omega_{\ell_1}, \ldots, \Omega_{\ell_L}]$ . In such case, the loudspeaker feed vector $\pmb{L}$ can be written as:
\begin{equation}
	\pmb{L} = \pmb{S}_n^m \pmb{D},
\end{equation}
where 
\begin{equation}
	\pmb{D} = \text{diag}(\alpha_n) [{\pmb{Y}_n^{m}(\Omega_{\ell_1})}^T, \ldots, {\pmb{Y}_n^{m}(\Omega_{\ell_L})}^T]
\end{equation}

is a $M \times L$ matrix known as the \textit{decoding matrix}, and $\text{diag}(\alpha_n)$ is a diagonal matrix of size $M$ containing the values of $\alpha_n$ along the main diagonal. 
$\pmb{D}$ is frequency-independent and depends solely on the loudspeaker array geometry. The reader is referred to [zotter? daniel? idhoa?] for more information about the vast field of study of ambisonic decoding.

\todo{figure from daniel? to sum up the subsection). Explain the allrad method and stuff because it will appear later on}.

@@

Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering [ambisonics in multichannel broadcasting and video]. 
By following this scheme, the four channels are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
\begin{equation}
	{\pmb{Y}_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
\end{equation}
with:
\begin{equation}
\begin{aligned}
	&W = Y_0^0(\Omega) \\
	&X = Y_1^1(\Omega) \\
	&Y = Y_1^{-1}(\Omega) \\
	&Z = Y_1^0(\Omega)
\end{aligned}	
\end{equation}

@@

@@

\section{SOFA}
 maybe as subsection of room acoustics

@@

\section{Practical Considerations}

>>>>

\chapter{Background}

\section{Conventions}
\subsection{Reference system}

@@

\subsection{Nomenclature}

Through the document, we refer to time-domain signals with lowercase, e.g. $x(t)$, with $t$ as the time index.

Time-domain signals transformed by the Short-Time Fourier Transform (STFT) are represented with uppercase, e.g. $X(k,n)$, where $k$ is the frequency bin index, and $n$ the time frame index.

Multichannel signals are in general denoted by a subscript variable index, usually with the letter $m$; for example, $x_m(t)$ or $X_m(k,n)$.
Signals with an integer subscript index, such as $x_0(t)$, represent a specific channel of the corresponding multichannel signal.

In the context of ambisonic, subscripts and superscripts are used in signal names with a specific meaning; check Section~\ref{sec:ambisonics} for a detailed explanation.

Vector notation is represented with boldface characters, e.g. $\bm{X}(k,n)$.  When used, the way to construct the vectors will be specified.

\todo{revise all vector notations}

@@

\section{Ambisonics}
\label{sec:ambisonics}

@@

The ordered set of values of all spherical harmonics up to order $N$, evaluated at the source position, is known as the \textit{ambisonic coefficients}:
\begin{equation}
	Y_n^m(\Omega_s) = [Y_0^0(\Omega_s), Y_1^{-1}(\Omega_s),  \ldots ,  Y_N^N(\Omega_s)]
	\label{eq:sphericalharmonicvector}
\end{equation}

Furthermore, the process of multiplying the signal $S$ by the ambisonic coefficients is known in the literature as the \textit{ambisonic encoding}. The resulting signal vector is usually referred to as the \textit{ambisonic} (or \textit{B-Format}) signal $\pmb{S}_n^m$:
\begin{equation}
	S_n^m = Y_n^m(\Omega_s) S
\end{equation}

@@

Let us consider a loudspeaker located at the angular position $\Omega_p$. The signal feed $P$ is \textit{decoded} from the ambisonic signal as:
\begin{equation}
	P = \sum_{n=0}^{N} \sum_{m=-n}^{n} Y_n^m(\Omega_s) S Y_n^m(\Omega_\ell)  \alpha_n 
	\label{eq:decoding}
\end{equation}

where $\alpha_n$ is a weighting factor which accounts for the beam directivity. Some of the weightings are widely used for its specific properties, such as \textit{max-rE} or \textit{in-phase}. Table~\ref{tab:alpha} shows the values of $\alpha_n$ for the different weightings, and Figure~\ref{fig:alpha} plots their first-order directive patterns.

\begin{table}[t]
\caption{Ambisonic decoding: standard values of $alpha_n$ weightings. Extracted from \cite{daniel2000representation}.}
\begin{center}
\begin{tabular}{cccccc}
\toprule
Decoding & $m$ & \multicolumn{4}{c}{$n$} \\ 
&  & 0 & 1 & 2 & 3 \\
\midrule
\textit{basic} & 0 & 1  \\
	& 1 & 1 & 1  \\
 	& 2 & 1 & 1 & 1 \\
 	& 3 & 1 & 1 & 1 & 1 \\
\midrule
\textit{max-rE} & 0 & 0.577  \\
	& 1 & 0.775 & 0.4  \\
 	& 2 & 0.861 & 0.612 & 0.305 \\
 	& 3 & 0.906 & 0.732 & 0.501 & 0.246 \\
\midrule
\textit{in-phase} & 0 & 0.333  \\
	& 1 & 0.5 & 0.1  \\
 	& 2 & 0.6 & 0.2 & 0.029 \\
 	& 3 & 0.667 & 0.286 & 0.071 & 0.008 \\
\bottomrule
\end{tabular}
\label{tab:alpha}
\end{center}
\end{table}

\begin{figure*}[htbp]
	\begin{center}
	\begin{minipage}[b]{0.9\linewidth}
		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/basic.png}}
	\end{minipage}
	\begin{minipage}[b]{0.9\linewidth}
		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/maxre.png}}
	\end{minipage}
		\begin{minipage}[c]{0.9\linewidth}
		\centerline{\includegraphics[width=\textwidth]{Figures/Introduction/inphase.png}}
	\end{minipage}
	\caption{Directive patterns of first-order ambisonic decoding.}
	\label{fig:alpha}
	\end{center}
\end{figure*}

The decoding equation \ref{eq:decoding} can be written in matrix form as:
\begin{equation}
	P = S_n^m {Y_n^m (\Omega_p)}^T \alpha_n
\end{equation}

where the superscript $T$ represents the matrix transposition. 
This equation can be extended to the usual case of decoding to a loudspeaker array, comprised of $L$ loudspeakers located at the positions $\Omega_L = [\Omega_{p_1}, \ldots, \Omega_{p_L}]$ . In such case, the loudspeaker feed vector $P_L$ can be written as:
\begin{equation}
	P_L = S_n^m D,
\label{eq:decodingequation}
\end{equation}
where 
\begin{equation}
	D = \text{diag}(\alpha_n) [{Y_n^{m}(\Omega_{p_1})}^T, \ldots, {Y_n^{m}(\Omega_{p_L})}^T]
\end{equation}

is a $M \times L$ matrix known as the \textit{decoding matrix}, and $\text{diag}(\alpha_n)$ is a diagonal matrix of size $M$ containing the values of $\alpha_n$ along the main diagonal. 
Although the matrix $D$ is frequency-independent and depends solely on the loudspeaker array geometry, in practical scenarios it is usual to include frequency-dependent weightings, $\alpha_n(k)$, to improve the broadband sound field reconstruction \cite{daniel2000representation}.

Furthermore, sound field reconstruction with Eq.~\ref{eq:decodingequation} is only possible when the loudspeakers are evenly located on the 3D space; in other words, the speaker layout must take the form of one of the five \textit{Platonic solids}: tetrahedron, cube, octahedron, dodecahedron or icosahedron.
Provided that this condition is usually difficult to fulfil in real scenarios, there are several methods which allow ambisonic decoding for such \textit{irregular} layouts. One of the most commonly used is the AllRAD method \todo{cite zotter}. AllRAD proposes a two step decoding: first, the ambisonic signal is decoded to a nearly-uniform layout of virtual speakers. Then, the signals of the virtual speakers are further distributed into the real speakers by the \textit{Vector-Based Amplitude Panning (VBAP)} method \todo{cite pulkki}.

% The reader is referred to [zotter? daniel? idhoa?] for more information about the vast field of study of ambisonic decoding.

@@

Historically, first-order ambisonic audio has followed what it might be called \textit{traditional B-Format} channel ordering [ambisonics in multichannel broadcasting and video]. 
By following this scheme, the four channels are referred to by the axis where the corresponding spherical harmonic steers, plus the name $W$ for the zeroth order component:
\begin{equation}
	{Y_n^m(\Omega)}^{\text{(FuMa CO)}} = [W, X, Y, Z]
\end{equation}
with:
\begin{equation}
\begin{aligned}
	&W = Y_0^0(\Omega) \\
	&X = Y_1^1(\Omega) \\
	&Y = Y_1^{-1}(\Omega) \\
	&Z = Y_1^0(\Omega)
\end{aligned}	
\end{equation}

@@

\todo{check here bold vector notation}

@@

ir, partes del ir

%%%%%%%%%%%%%% cogido del paper de MMSP
Reverberation time $T_{60}$ \cite{kuttruff2016room} might be one of the most widespread acoustic parameters; it represents the time required 
%for a sound in the diffuse field to decrease its energy envelope by 60 dB. 
for the reverberant sound field power to decay by 60 dB.
Reverberation time can be accurately computed from the room geometry \cite{sabine1927collected} or from the Impulse Response (IR) \cite{schroeder1965new}; the problem of $T_{60}$ estimation just from observations of the reverberant signal itself is referred to as the blind reverberation time estimation, and it still remains an open research question.

% TODO
reverberation time, sabine, schroeder
%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{SOFA}
 maybe as subsection of room acoustics

@@

\section{Signal Model}

Let us consider a sound source represented by the signal $s(t)$, located in a given acoustic enclosure. Assuming that the system is Linear and Time-Invariant (LTI), it can be represented by its room impulse response $h(t)$. The resulting reverberant signal $x(t)$ can be therefore described as the \textit{convolutive mixture} of the source and the RIR: 
\begin{equation}
	x(t) = s(t) \ast h(t).
\label{eq:convolutivemixture}
\end{equation}

When dealing with multichannel room impulse responses, as it is the case in ambisonics, the multichannel reverberant signal $x_m(t)$ is obtained by the convolutive mixture of each RIR channel independently:
\begin{equation}
	x_m(t) = s(t) \ast h_m(t).
\label{eq:convolutivemixturemultichannel}
\end{equation}

The time domain convolution operation, under certain assumptions, is equivalent to the multiplication in frequency domain. By doing so, Eq.~\ref{eq:convolutivemixture} can be expressed as:
\begin{equation}
	X(k,n) = S(k,n) H(k,n).
\label{eq:multiplicativetransferfunction}
\end{equation}

Eq.~\ref{eq:multiplicativetransferfunction}, also known as the \textit{Multiplicative Transfer Function (MTF) model} is only valid when the length of the filter $h(t)$ is smaller than the length of the analysis window used in the STFT.

On the contrary, when the filter $h(t)$ spans across several analysis windows, the resulting model is referred to as the \textit{Convolutive Transfer Function (CTF) model}:
\begin{equation}
	X(k, n) = \sum_{l=0}^{L_h-1} H(k, l) S(k, n-l), 
\label{eq:convolutivetransferfunction}
\end{equation} 
where $L_h$ is the length of the filter $H(k, n)$ in time frames.

\todo{isotropic noise}

\section{Practical Considerations}
% TODO I don't remember what is this

### 2020-06-01 manuscript.tex
updated
update 20200602

%COHERENCE ESTIMATION
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{latexsym}% 
\usepackage{lineno}   
\usepackage{array}
\usepackage{adjustbox}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{newpxtext} 
\usepackage{cite}
\usepackage{booktabs}

@@

@@

%ADD YOUR DATA
\title{Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}}
\subtitle{\todo{subtitle Required?}}
\author{Author: Andrés Pérez López}
\thyear{2020}
\department{of Information and Communication Technologies}
\supervisor{Dr. Emilia Gómez Gutiérrez\\Dr. Adan Garriga Torres}

@@

%INSERT CHAPTERS
\input{Chapters/Introduction.tex}
\input{Chapters/ScientificBackground.tex}
\input{Chapters/CoherenceEstimation}
\input{Chapters/AutoregressiveReverbEstimation}
\input{Chapters/SELD}
\input{Chapters/DataGeneration}
\input{Chapters/Conclusions}

%ADD BIBLIOGRAPHY
\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib}

>>>>

%COHERENCE ESTIMATION
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{latexsym}% 
\usepackage{lineno}   
\usepackage{array}
\usepackage{adjustbox}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{newpxtext} 
\usepackage{cite}
\usepackage{booktabs}
%\usepackage{hyperref}
%\hypersetup{pdftex,colorlinks=true,allcolors=blue}
%\usepackage{hypcap}

@@

%\usepackage{hyperref}

@@

%ADD YOUR DATA
\title{Spherical harmonic domain methods for 3D audio production enhancement \todo{(check capitalization)}}
\subtitle{\todo{subtitle Required?}}
\author{Author: Andrés Pérez López}
\thyear{2020}
\department{of Information and Communication Technologies}
\supervisor{Dr. Adan Garriga Torres\\Dr. Emilia Gómez Gutiérrez}

@@

%INSERT CHAPTERS
\input{Chapters/Introduction.tex}
\input{Chapters/ScientificBackground.tex}
\input{Chapters/CoherenceEstimation}
\input{Chapters/ReververationTimeEstimation}
\input{Chapters/SELD}
\input{Chapters/DataGeneration}
\input{Chapters/Conclusions}

%ADD BIBLIOGRAPHY
\bibliography{Bibliography/CoherenceEstimation, Bibliography/SELD.bib, Bibliography/Ambiscaper.bib, Bibliography/AmbisonicsDRIR.bib, Bibliography/ReverberationTimeEstimation.bib}
